\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-16 trunk/concepts.tex
merged minor mods to europa secs

\section{Foundational Concepts}
\label{sec:concepts}

Our system, \texttt{T-REX} (or the Teleo-Reactive EXecutive) is an
adaptive, artificial intelligence based controller and provides
general framework for building reasoning systems for real-world
autonomous vehicles. At MBARI \texttt{T-REX} is used for AUV control;
another instantiation of the system is being used for control of a
terrestrial personal robot \cite{pr2, Meeussen:2010dn}. The development of
\texttt{T-REX} has been targeted at surveying a number of
oceanographic features which are dynamic and unpredictable
spatio-temporally. Typically this requires our AUVs to balance the
goals of coverage spatially while opportunistically following features
of scientific interest and to do so while being operationally aware of
its own limitations in terms of resources (typically the battery state
of charge) and overall proximity to other observational assets for
obtaining scientific ground-truth. We are therefore interested in
representational frameworks that allow robots to pursue long-term
objectives while choosing short-term gain and can gracefully deal with
exogenous or endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use a
temporal constraint-based planner EUROPA$_2$ with a demonstrated
legacy of having flown on NASA space missions
\cite{Jonsson00,bresina05}. Our autonomy architecture brings three key
innovations for AUV adaptation: the use of flexible plan
representations, compositional control with the use of partitioned
networks and off-line learning to inform insitu environmental state
estimation. In this section we detail some of the key concepts
relevant to flexible plan representation. Section \ref{sec:arch} has
details the \texttt{T-REX} partitioned architecture and estimation
driven vehicle adaptation.

% \texttt{T-REX} uses EUROPA$_2$ a constraint-based temporal planner
% \cite{Jonsson00,Frank2003} developed for NASA missions, most recently
% for the Mars Exploration Rovers \cite{bresina05}.

EUROPA$_2$ uses a domain model written in a declarative language
(NDDL: New Domain Description Language), together with initial
conditions and goals also specified in NDDL, to construct a set of
temporal relations that must be true at the start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, EUROPA$_2$ can select
a set of conditions that should be true in the future, where some of
these conditions will correspond to actions the agent must take. The
planner can backtrack and try another path during search if a goal
cannot be reached while being capable of discarding unachievable
goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,Alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

EUROPA$_2$ uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a EUROPA$_2$ domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} relations shown above.}
\label{fig:allen-algebra}
\vskip-0.3cm
\end{figure}

Each of the token variables, including the parameter variables, has a
domain of values assigned to it.  The variables may also participate
in constraints that specify which value combinations are valid.  To
allow the specification of multiple values, e.g, to express a range of
possible start times, variables are used to specify parameter, start
and end time values for a token.  As a result, a token $T$ is a tuple
$\langle v, P, s, e \rangle$, where $v$ is a variable denoting a state
variable, $P$ is the name of a procedure and $s$ and $e$ (satisfying
$s\leq e$) are variables encoded as flexible token start and end time
points. Consequently, partial plans represent a \textit{range of
  possible solutions} rather than an explicit definition of a single
trajectory of state.  Fig. \ref{fig:flex-timelines} shows an example
of a part of such plans with tokens representing constraints on
flexible temporal intervals. Systematic computation is reflected by
the use of the \texttt{Allen Algebra} \cite{allen84} temporal
relations shown in Fig. \ref{fig:allen-algebra}. It is the use of such
interval arithmetic in the EUROPA$_2$ domain model that allows
encoding of temporal constraint relationships between tokens on the
same or other concurrent timelines. Details of propagation algorithms
or of interval logic are beyond the scope of this paper.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/FSM-transition.pdf}
\caption{\small Timeline evolution can be described by simple finite
  state transitions where temporal relationships describe the arc
  transition and temporal extant of state are token
  durations. Constraints here are for \emph{meets} or its equivalent
  inverse \emph{met\_by}.}
\label{fig:FSM}
\vskip-0.3cm
\end{figure}

In describing the evolution of state via timelines, transitions within
a timeline can be described by simple finite state
machines. Fig. \ref{fig:FSM} for example describes a timeline
evolution for an abstract timeline entity describing navigation. Arc
transitions represent temporal constraints and token duration are
represented by temporal extant of state. The entire vehicle can then
be modeled as a number of state transitions with concurrency for each
of the modeled state variables defined a priori. However this simple
view is often complicated when temporal constraints between state
variables are articulated in our domain models. If there is a
criticism of this approach, modeling then, requires a proper balance
between abstraction, state variable description and concurrency
described by the inter-connectedness between such finite state
machines.

\begin{figure*}
\centering 
\subfloat[]{\label{fig:plan-evolve1}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-1.pdf}} 
\subfloat[]{\label{fig:plan-evolve2}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-2.pdf}} 
\subfloat[]{\label{fig:plan-evolve3}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-3.pdf}} 
\caption{\small An illustrative plan synthesis example with concurrent
  timelines. Each timeline is an instantiation of a subsystem tracked
  by the planner. Tokens describe the instantiated state of subsystem
  at a particular time instance and enforce temporal and parametric
  constraints. Abstract goals are decomoposed into successively less
  abstract tokens and instantiated in co-temporal timelines using
  \texttt{Allen Algebra} relations. All tokens represent flexible
  start/end times. \ref{fig:plan-evolve1} shows an initial state
  evolving into \ref{fig:plan-evolve2} and \ref{fig:plan-evolve3}.}
  \label{fig:Plan-evolve}
  \vskip-5pt
\end{figure*}

Fig. \ref{fig:Plan-evolve} shows an illustrative example. We show four
timelines which track \texttt{Goal, Path, Navigation} and
\texttt{Command} state over time. Tokens on the goal timeline indicate
a survey within a bounded volume and a min/max depth envelope. As the
Volume Survey sub-goals on the \texttt{Path} timeline, the initial
token on that timeline gets \emph{squeezed} to have dependencies of
the goal token instantiated. Initially this dependency are the tokens
\texttt{Go(x1,y1)} and \texttt{Go(xn,yn)} indicative of the area of
coverage. The first of these in turn generates a sub-goal on the same
timeline to go to the next intermediate waypoint, which is illustrated
in Fig. \ref{fig:plan-evolve2}. Subsequent sub-goals (not shown in the
figure) ultimately generate all the tokens to completion for this
timeline. Fig. \ref {fig:plan-evolve3} shows a snapshot further along
in plan generation and shows additional subgoals on the \texttt{Path,
  Navigation} and \texttt{Command} timelines. Each token can thus
trace its causality when instantiated in the plan. The domain model
provides source of these temporal and parametric dependencies in the
form of rules.

\begin{minipage}[c]{\textwidth}

% \framebox[\textwidth][t]{
Volume\_Survey(x$_1$,y$_1$,x$_n$,y$_n$,Min\_Depth, Max\_Depth)\\
$\Rightarrow$ \\
met\_by Go(x$_1$,y$_1$,Min\_Depth); \\
starts [50,0] Go(x$_2$,y$_2$,Min\_Depth,Max\_Depth); \\
\ldots{} \\
ends [50,0] Go(x$_{n-1}$,y$_{n-1}$, Min\_Depth,Max\_Depth); \\
meets Go(x$_n$,y$_n$,Min\_Depth, Max\_Depth); 
\end{minipage}

>>>>

\subsubsection{\eu Plan Representation}
\label{sec:europa:pr}

\eu follows the representation outlined by the CAIP framework, it also adds a few elements to it that are necessary for an efficient implementation of constraint-based planning. Below is a brief summary of the main elements, followed by a detailed description for each one:

\begin{enumerate}
\item \textbf{Tokens} Temporally scoped entities that correspond to Intervals in the CAIP framework, are used to represent  state and actions in time. A\textbf{Token State model} is defined to support an efficient implementation of plan-space search algorithms.

\item \textbf{Domains, Variables and Constraint} are used to represent restrictions in the problem domain in terms of a CP formulation.

\item \textbf{Domain Rules} are used to describe internal and external relationships on and between tokens, as presecribed by the CAIP framework.

\item \textbf{Objects} are used to represent problem domains in a more modular and scaleable way. \textbf{Timelines and Resources} are built in implementations of Object types that appear frequently in real life domains.

\end{enumerate}

\begin{description}

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used.  As
  explained in the \textsf{CP} section, a variable can take values
  from a domain. In \eu, domains are defined over a specific data
  type, the primitive data types supported are:

  \begin{enumerate}
  \item \textit{String}: sequences of characters, like "Red",
    "Spirit", etc

  \item \textit{Boolean}: {True,False}

  \item \textit{Numeric}: integers and floats

  \item \textit{Object}: reference to an object instance (see below
    for a description of objects in \eu)

\end {enumerate}

Two types of variable Domains can be defined over a data type:

  \begin{enumerate}
  \item \textit{Enumerated}: Finite sets over any data type, specified
    explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
  \item \textit{Interval}: Only for Numeric data types, they are
    specified by a [LowerBound,UpperBound] pair, for instance: [1,10],
    [5.0,100.0], etc.
  \end {enumerate}

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans. \comment{say more about the benefits of flexibility?}

\item[\textbf{Constraints}] valid plans in most problem domains have
  to satisfy a number of restrictions, for instance, a Shopping Agent
  may only be able to perform one task at a time, may have limited
  time or an energy or fuel budget to perform a task, and so on. In
  \eu, these restrictions are represented by Constraints, which can be
  defined over any combination of Variables in a domain. \eu provides
  a Constraint Library that already implements many useful constraints
  (temporal, resource, relational, etc), it is also possible to add
  new constraints if required by a particular domain.

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered Objects. As in the case with object-oriented
  analysis and design, one can seek out the nouns in any domain
  description to find likely objects. In the Shopping Agent example,
  we might consider the Agent itself, Products and Locations all to be
  objects. Objects have state and behavior. For example, a Shopping
  Agent can have:

  \begin{itemize}

  \item State: its location, a bag that contains the products it has
    already purchased (the bag can in turn be another object), etc.

  \item Behavior: go to a location to look for a product, perform a
    purchase, return home, etc.

  \end{itemize}

  Objects that have similar state and behavior can be described
  generically in terms of object types or classes. \eu allows the
  definition of classes in the same way that it is done in popular
  object-oriented programming languages, where a class can have
  attributes (represented by Variables), and classes can be arranged
  in a single-parent hierarchy through inheritance.  However, one key
  aspect needs to be introduced to class definition in \eu that is not
  found in object-oriented programming: to describe state and behavior
  for the purposes of planning we need to build on the formalism of
  first order logic as explained below.

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \comment{citation?}.  In
  \eu, we define such relations between variables whose domains are
  sets of objects and sets of properties to describe state and
  behavior. For example, we might use a predicate \texttt{At($a$,$l$)}
  to indicate that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking acting to
  buy product $p$. Note that $a$ is a variable which may have a number
  of possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value in them, or grounded plans, where single values will
  be specified for each variable as we saw in the case of the
  N-queen's problem.

  In general, to come up with an executable plan it is not sufficient
  to state predicates that describe required state or behavior without
  also specifying some temporal extent over which each of those
  predicates hold. Predicates that are always true can be thought to
  hold from the beginning to the end of time. However, in practice,
  the temporal extent of interest must be defined with timepoints to
  represent its start and end. So we might want to write
  \texttt{At($a$,$s$,$e$,$l$)} to indicate that the agent $a$ is at
  location $p$ from time $s$ to time $e$. In fact, this pattern of
  using such predicates to describe both state and behavior of objects
  is so prevalent in \eu that we have introduced a special construct
  called a Token which has the built in variables to indicate the
  object to which the statement principally applies and the timepoints
  over which it holds. A Token is an instance of a predicate that
  represents and object's state or behavior and is defined over a
  temporal extent. In \eu, all predicate instances are Tokens. Also,
  in the same way that objects are described generically by classes,
  in \eu Tokens can be described generically by Token Types. Every
  token has five built-in variables:

  \begin{enumerate}

  \item \textit {start}: The beginning of the temporal extent over
    which the predicate is defined.

  \item \textit {end}: The end of the temporal extent over which the
    predicate is defined.

  \item \textit {duration}: The duration of the temporal extent over
    which the predicate is defined. The constraint \textit{start} $+$
    \textit{duration} = \textit{end} is enforced automatically.

  \item \textit{object}: The set of objects to which a token might
    apply. In a grounded plan each Token applies to a specific Object,
    reflecting the intuition that we are using Tokens to describe some
    aspect of an Object (i.e. its state or behavior) in time. However,
    in a partial plan, the commitment to a specific object may not yet
    have been made.

  \item \textit{state}: Tokens can be \texttt{ACTIVE, INACTIVE,
      MERGED}, or \texttt{REJECTED}. The state variable captures the
    token's current state and its reachable states through further
    restriction.  This is required to support CAIP's approach to
    planning where goals and subgoals can be satisfied by either
    adding new intervals (Tokens) to the plan, or by merging with
    existing intervals. Token State lifecycle is examined in detailed
    in a separate section below.

  \end{enumerate}

\item[\textbf{Built-in Object Types}] In a domain model, there may be
  some classes that don't have any time-dependent state or behavior,
  and therefore those classes will not have any Token types associated
  with them, for instance in the Shopping Agent model, the set
  locations and products are static for the agent's purposes so we may
  only want to say what they are but not have any state or behavior
  associated with them.

  However, the most common case in any non-trivial domain model is
  that Objects will have many associated Tokens in order to describe
  their state and behavior throughout a plan. There are a couple of
  Object Types that are so commonly used in domain descriptions, that
  \eu provides a built-in implementation for them:

\begin{enumerate}

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{add small
    example?}

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are such a common requirement for
  \eu users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided for
  them. Instances of classes derived from a Resource will induce
  ordering requirements on their Tokens in order to ensure that the
  level of the resource remains within specified limits.  \comment{add
    small example?}
\end{enumerate}

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  Tokens.  Goals can be posted explicitly, for instance, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  compatibilities (described in the CAIP framework above), for
  instance, when the Shopping Agent must get to a Location where a
  Product is available before it is able to purchase it, in this case
  the original goal of purchasing the product will spawn a subgoal of
  being at a specific location.

  Looking at a how a simple Shopping Agent goal would be addresed by
  \eu helps illustrate how Token States support the planning process.

  Let's assume we start with a partial plan that places the Shopping
  agent at home at time 0 and with no Products in its possession, then
  we post a goal for the Shopping Agent to own a Drill at time 10.

  \comment{Add figure showing an AgentLocation timeline with an
    At(Home) token with start=0 and end open, an empty AgentBag
    timeline, and a Own(Drill) token with start=10 floating below both
    timelines}

Notice that the \texttt{Own(Drill)} token that represents the goal is
not immediately placed on the \texttt{AgentBag} timeline since \eu
must determine whether it can achieve that goal and how. When the goal
is posted and before any decisions are made on it, the
\texttt{Own(Drill)} token is created in an INACTIVE state, then \eu
looks at the current partial plan and considers 2 alternatives:

\begin{enumerate}

\item If there are any tokens in the plan that are compatible with the
  new goal, \eu may attempt to satisfy the goal by merging it with any
  of them (if there is more than one it can try them one at a time),
  in this case, the state of the token would change to MERGED.

\item Instead of merging with existing tokens in the plan, \eu may
  decide to insert the new token into the plan, in that case its state
  would change to ACTIVE.

\end{enumerate}

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible with
the inactive token. Two tokens are merge-compatible if they are
instances of the same predicate and that no intersections between
corresponding variables are empty. The effect of merging is
illustrated in Figure 3.

\comment{Add figure showing the effect of merging; all figs on this
  section are conjoined. JB to sketch it out.}

When a token is activated all compatibilities associated with its
token type are evaluated, which may lead to new subgoals being
generated. In our example, there should be a compatibility stating
that the Shopping Agent must be at the same location as the product it
is purchasing. In that case, activating the \texttt{Own(Drill)} token
will cause a new \texttt{At(Drill.location)} to be created in an
INACTIVE state, which in turn will have to be dealt with.

\comment{Add figure showing the effect of activating the token by
  putting it on the AgentBag timeline, and a new At(Drill.location)
  token floating below}

If neither activating nor merging works, a token may become REJECTED,
this is possible only for goals posted by the end user and specified
as optional. Subgoals generated as a result of token activation cannot
be rejected, if a subgoal cannot be satisfied through activation or
merging then it will cause the original activation decision that
spawned it to be retracted (\eu's search mechanisms are explained in
detail in a separate section later on).

As we can see, a token's state can go from INACTIVE to ACTIVE, MERGED
or REJECTED, this lifecycle is illustrated in the figure below.

\comment{Add figure showing possible state transitions. Need a FSM.}

\item[\textbf{Rules}] In order for a plan to be valid, it must comply
  with all rules and regulations pertinent to the application domain
  in question. Rules govern the internal and external relationships of
  a token. For example, consider a parameterized action
  \texttt{Go(origin,destination)} which describes a Shopping Agent
  traveling from one location ( \textit{origin}) to another (
  \textit{destination}). The parameters are instantiated on a token as
  variables whose domain of values is the set of all locations in a
  given problem. A rule governing an internal relationship among token
  variables might stipulate that a transition must involve a change in
  location. This can be easily expressed as a constraint on the
  definition of a predicate of the form: \textit{origin !=
    destination}. It makes sense to further stipulate that the agent
  must be located at \textit{origin} before travel can happen and the
  agent must be located at \textit{destination} when completed. This
  is an example of a rule governing an external relationship among
  tokens. It specifies a requirement that tokens of the predicate
  \texttt{At} that represent the location state for the Shopping Agent
  precede and succeed tokens of the predicate \texttt{Go} that
  represents traveling.

  \comment{ Add figure showing internal and external relationships
    from rules on action Go. Need a fig from JB}

Figure xx illustrates the entities and relations involved in
specifying such a rule on a \texttt{Go} action. The token on which the
rule applies is referred to (\texttt{Go}) as the master. Each
\texttt{At} token required by the master is referred to as a
slave. All variables are indicated by name and their domains are
expressed as intervals in the case of temporal variables and as
enumerations for the remainder. Application of a rule on a token can
thus cause slave tokens, variables, and constraints to be introduced.

The capability of a domain rule to cause a slave token to be created
is a key vehicle through which planning occurs. Semantically, this
rule imposes a requirement for supporting (slave) tokens to be in the
plan in order for the master to be valid.

\comment {talk about condition/effect annotations?}

\end{description}

%\begin{figure*}
%\centering 
%\subfloat[]{\label{fig:plan-evolve1}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-1.pdf}} 
%\subfloat[]{\label{fig:plan-evolve2}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-2.pdf}} 
%\subfloat[]{\label{fig:plan-evolve3}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-3.pdf}} 
%\caption{\small An illustrative plan synthesis example with concurrent
%  timelines. Each timeline is an instantiation of a subsystem tracked
%  by the planner. Tokens describe the instantiated state of subsystem
%  at a particular time instance and enforce temporal and parametric
%  constraints. Abstract goals are decomoposed into successively less
%  abstract tokens and instantiated in co-temporal timelines using
%  \texttt{Allen Algebra} relations. All tokens represent flexible
%  start/end times. \ref{fig:plan-evolve1} shows an initial state
%  evolving into \ref{fig:plan-evolve2} and \ref{fig:plan-evolve3}.}
%  \label{fig:Plan-evolve}
%  \vskip-5pt
%\end{figure*}

%Fig. \ref{fig:Plan-evolve} shows an illustrative example. We show four
%timelines which track \texttt{Goal, Path, Navigation} and
%\texttt{Command} state over time. Tokens on the goal timeline indicate
%a survey within a bounded volume and a min/max depth envelope. As the
%Volume Survey sub-goals on the \texttt{Path} timeline, the initial
%token on that timeline gets \emph{squeezed} to have dependencies of
%the goal token instantiated. Initially this dependency are the tokens
%\texttt{Go(x1,y1)} and \texttt{Go(xn,yn)} indicative of the area of
%coverage. The first of these in turn generates a sub-goal on the same
%timeline to go to the next intermediate waypoint, which is illustrated
%in Fig. \ref{fig:plan-evolve2}. Subsequent sub-goals (not shown in the
%figure) ultimately generate all the tokens to completion for this
%timeline. Fig. \ref {fig:plan-evolve3} shows a snapshot further along
%in plan generation and shows additional subgoals on the \texttt{Path,
%  Navigation} and \texttt{Command} timelines. Each token can thus
%trace its causality when instantiated in the plan. The domain model
%provides source of these temporal and parametric dependencies in the
%form of rules.

%\begin{minipage}[c]{\textwidth}
%\vspace{+0.5cm}
% \framebox[\textwidth][t]{
%Volume\_Survey(x$_1$,y$_1$,x$_n$,y$_n$,Min\_Depth, Max\_Depth)\\
%$\Rightarrow$ \\
%met\_by Go(x$_1$,y$_1$,Min\_Depth); \\
%starts [50,0] Go(x$_2$,y$_2$,Min\_Depth,Max\_Depth); \\
%\ldots{} \\
%ends [50,0] Go(x$_{n-1}$,y$_{n-1}$, Min\_Depth,Max\_Depth); \\
%meets Go(x$_n$,y$_n$,Min\_Depth, Max\_Depth); 
%\end{minipage}

% \end{enumerate}

### 2012-02-16 trunk/europa-basics.tex
merged minor mods to europa secs

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \comment{citation} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. \comment{A figure of a
  constraint network preferably of the shopping example would be
  useful here.}

@@

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
mutual exclusion and concurrency using propositions \comment{citation
  needed}. 
Constraint Programming is a framework that lends itself very well to represent and reason about all of those elements \comment{Ixtet citation?},
consequently, there has been a good amount of work building automated planners that take advantage of CP techniques.
The most common approach in constraint-based planning has been to encode the entire planning problem as a CP problem, 
where action choices and relationships are represented as variables and/or constraints \comment {citations 158,519,539,556 from malik's book}.
The advantage of using a pure CP formulation is that standard CP solvers can be then used to solve the planning problem, however, this approach also has a couple of significant drawbacks:

>>>>

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \cite{Bruckera99} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. \comment{A figure of a
  constraint network preferably of the shopping example would be
  useful here. JB to generate a napkin fig}

@@

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,

% mutual exclusion and concurrency using propositions
% \cite{ghallab04}. While CSP representations have traditionally had an
% edge in formulating and representing planning problems,
% % It is straightforward to represent and reason about all of
% % those elements using a CSP representation, as a result there has been
% % some work on doing automated planning while taking advantage of CSP
% % (TODO: ref). Traditionally, the entire planning problem is translated
% % into a CSP and then solved using traditional CSP methods, this
% leading to a formulation where action choices and relationships are
% represented as variables and/or constraints, there are at least two
% major drawbacks:

mutual exclusion and concurrency using propositions
\comment{ghallab04}.  Constraint Programming is a framework that lends
itself very well to represent and reason about all of those elements
\comment{Ixtet citation?}, consequently, there has been a good amount
of work building automated planners that take advantage of CP
techniques.  The most common approach in constraint-based planning has
been to encode the entire planning problem as a CP problem, where
action choices and relationships are represented as variables and/or
constraints \comment {citations 158,519,539,556 from malik's book}.
The advantage of using a pure CP formulation is that standard CP
solvers can be then used to solve the planning problem, however, this
approach also has a couple of significant drawbacks:
